---
title: "Subs-predict - initial stuff to get started"
date: "March 16, 2018"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=F, message=F)

library(dplyr)
library(ggplot2)
library(caret)
library(dplyr)
library(DT)
library(lubridate)

source('aggregation.R')

# some configuration
SLIMDATA <- T
slimfactor <- 1.0

periodendcutoff <- parse_date_time('2017-01-01', 'ymd')
beginwindow <- parse_date_time('2017-09-01', 'ymd')
endwindow <- parse_date_time('2018-01-01', 'ymd')


library(lubridate)
```

Getting the data

```{r reading, tidy=F}
if (!exists('customersorg'))
{
  # don't read again, we've already got it
  customersorg <- read.table(file = 'data/CustomersV2.export.tsv', sep = '\t', header = TRUE, stringsAsFactors = F)
  gdp <- read.table(file = 'data/gdp-simple.export.tsv', sep = '\t', header = TRUE, stringsAsFactors=F)
  colnames(gdp)<-c('marketname', 'gdppercapita')
  subsorg <- read.table(file = 'data/SubscriptionsV2.export.tsv', sep = '\t', header = TRUE,
                     stringsAsFactors=F)
}
```

Optionally slimming the data for faster development

```{r slimming, tidy=F}
if (SLIMDATA)
{
  set.seed(1)
  index <- caret::createDataPartition(customersorg$customerid, p = slimfactor, list = FALSE, times = 1)
  customers <- customersorg[index,]
  custids <- customers %>% dplyr::select(customerid)
  subs <- subsorg %>% dplyr::inner_join(custids, by = 'customerid')
} else {
  customers <- customersorg
  subs <- subsorg
}
```

Fixing the dates
```{r dates, tidy=F}
subs$periodend<-as.Date(subs$periodend)
subs$startmonth<-as.Date(subs$startmonth)
subs$endmonth<-as.Date(subs$endmonth)
subs$startmonth<-ymd(subs$startmonth)
subs$endmonth<-ymd(subs$endmonth)
# Create the months variable
subs$months=(year(subs$endmonth)-year(subs$startmonth))*12+
  (month(subs$endmonth)-month(subs$startmonth))
```

Filtering the data
leaving only those with subscription period 1,3,12,24

```{r filtering, tidy=F}
x1 <- unique(subs$customerid[!(subs$months %in% c(1,3,12,24))])
# period starts after march 2018
x2 <- unique(subs$customerid[subs$startmonth>'2018-03-01']) 
x3<-unique(c(x1,x2))
rm(x1);rm(x2)
# Subsetting the data
subs1<- subs[!(subs$customerid%in% x3),]
```

Removing no need variables and preparing a bit

```{r, tidy=F}
# Removing variables
subs1$periodstart<-NULL
subs1$periodend<-NULL
subs1$billingcurrency<-factor(subs1$billingcurrency)
```

Data preperation
Here we create 3 new variables:
lag: The previous state, is 0 if is new customer
months_total: How many months do we have of total subscription
times: How many times he bought subscription
```{r dplyr}
## Sorting by ID and date
subs1$One<-1 ## We need this for later
subs1$months1<-subs1$months
subs1$months<-as.character(subs1$months)
#subs2<-subs1[1:25000,]
subs1<-subs1 %>%
  
  arrange(customerid, startmonth) %>%
  group_by(customerid) %>%
  # Take the last row for each customerid, duplicate it
  filter(row_number()==n()) %>%
  #mutate(months = as.character(months)) %>%
  mutate(months = ifelse(endmonth<'2018-03-01', 'churn', 'active')) %>%
  #We need this for sorting later
  mutate(subscriptionid=subscriptionid+0.001)%>%
  bind_rows(subs1)%>%
  arrange(customerid, subscriptionid) %>%
  # Total months and number of subscriptions before the given state/event
  mutate(months_total=lag(cumsum(months1), n=1, default=0), times = dplyr::lag(cumsum(One), n=1, default=0)) %>%
  # previous subs type, 0, if this if first
  mutate(lag = dplyr::lag(months, n=1, default='0'))

subs1 <- subs1 %>%
  mutate(lead = dplyr::lead(months, n=1))

subs1$months1<-NULL
subs1$One<-NULL 

# get rid of dplyr dataframe structure
subs1<-as.data.frame(subs1)
# take out actives
subs1<-subs1[subs1$months !='active',]
```

Data preparation and joining with customerid and gdp

```{r joining}
subs1$months<-factor(subs1$months, levels=c('1','3','12', '24', 'churn'), 
                     labels=c('1','3','12', '24',  'churn'))  
subs1$lag<-factor(subs1$lag, levels = c('0','1','3','12', '24'), 
                  labels = c(c('0','1','3','12', '24')))
subs1$lead<-factor(subs1$lead, levels = c('churn','1','3','12', '24'), 
                  labels = c(c('churn','1','3','12', '24')))

# joining with customers
subs2<-plyr::join(subs1, customers, by='customerid', type='left', match='all')
# Joining with gdp data
subs2<-plyr::join(subs2, gdp, by='marketname', type='left', match='all')
```

Changing the types of data to categorical when needed and scaling the numeric variables

```{r var_types}
subs2$isfreemium<-factor(subs2$isfreemium, levels=c(0,1), labels=c('No', 'Yes'))
subs2$segment<-factor(subs2$segment)
subs2$productversion<-factor(subs2$productversion)
subs2$isquickpurchase<-factor(subs2$isquickpurchase, levels=c(0,1), labels=c('No', 'Yes'))
subs2$firstdevice<-factor(subs2$firstdevice)
subs2$marketname <- factor(subs2$marketname)
subs2$model31224 <- factor(subs2$model31224)
subs2$channelcat <- factor(subs2$channelcat)
subs2$siteverkey <- factor(subs2$siteverkey)
subs2$currency <- factor(subs2$currency)
## Scale gdp data
gdpsc<-as.vector(scale(subs2$gdppercapita, center=T, scale=T))
subs2$gdppercapita<-gdpsc
```

Slim to only recent expiries

```{r}
  subs3 <- subs2 %>% dplyr::filter(endmonth >= periodendcutoff)
```

Add churncat and other cats to subs
```{r}

tmp <- subs3 %>%
  addSiteVerCats() %>%
  addAgeAtStartCatsChurn() %>%
  addSiteVerKeyCatChurn()



subs4 <- subs3 %>% 
  addChurnCatsEtc()
```

```{r}
summary(subs4)
```

Prepare churntable that we want to predicdt

```{r}
churntable <- subs4 %>%
  # remove synthetic "next" event
  dplyr::filter(months != 'churn') %>%
  # restrict to a recent expiry window
  dplyr::filter(endmonth >= beginwindow & endmonth < endwindow) %>%
  dplyr::mutate(churnind = ifelse(lead == 'churn', 1, 0)) %>%
  dplyr::group_by(siteverkeycatchurn, marketcat, months, ageatstartcatchurn, custchosensubslengthcatchurn, churncatnomarket) %>%
  dplyr::summarise(numobs = length(customerid), churned = sum(churnind)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(churn = churned / numobs,
                renewrate = 1 - churn,
                monthchurn = 1 - renewrate ^ (1/as.double(months)),
                logmonthchurn = log(monthchurn),
                weight = numobs / sum(numobs)) %>%
  # NB! Does this introduce a bad bias ????
  dplyr::filter(churn != 0)
```

Calculate model

```{r}
new_model=glm(logmonthchurn~marketcat+churncatnomarket, data=churntable, weights = churntable$weight)
summary(new_model)

anova(new_model, test="Chisq")
```

What is the dimensionality?
```{r}
dimstructure <- predictions %>% dplyr::group_by(churncatnomarket) %>% dplyr::summarise()
length(dimstructure$churncatnomarket)
dimmarket <- predictions %>% dplyr::group_by(marketcat) %>% dplyr::summarise()
length(dimmarket$marketcat)
length(dimstructure$churncatnomarket) * length(dimmarket$marketcat)
```



Add predictions

```{r}
predictions <- churntable
predictions$predictlogmonthchurn <- predict(new_model, type="response")
predictions <- predictions %>% dplyr::mutate(predictmonthchurn = exp(predictlogmonthchurn)) 
```

Print the predictions

```{r}
tblpred <- predictions %>%
  dplyr::select(siteverkeycatchurn, marketcat, ageatstartcatchurn, custchosensubslengthcatchurn, months, 
                churncatnomarket, monthchurn, predictmonthchurn) %>%
  dplyr::arrange(siteverkeycatchurn, marketcat, ageatstartcatchurn, custchosensubslengthcatchurn, months, churncatnomarket)
DT::datatable(tblpred, options = list(pageLength = 50), rownames = FALSE, filter = 'top') %>%formatRound( columns=c(6,7), digits=3)

```



Do some plotting

```{r}
plot3mhistory <- function(market, pred)
{
  tmp <- pred %>%
    dplyr::filter(marketcat == market &
                    (
                      (ageatstartcatchurn == 0 & months == 1 & custchosensubslengthcatchurn == 3) |
                      (ageatstartcatchurn != 0 & months == 3 & custchosensubslengthcatchurn == 'gen')
                    ) &
                    siteverkeycatchurn != "ORG"
                  ) %>%
    dplyr::select(ageatstartcatchurn, monthchurn, predictmonthchurn) %>%
    tidyr::gather(key = variable, value = value, -ageatstartcatchurn) %>%
    dplyr::arrange(ageatstartcatchurn)
  
  tmp %>%
    ggplot(aes(x=ageatstartcatchurn, y=value, color=variable)) + geom_line() + geom_point() + ggtitle(market)
}

plot3mhistory("DK", predictions)
plot3mhistory("NO", predictions)
plot3mhistory("SE", predictions)
plot3mhistory("US", predictions)
plot3mhistory("FR", predictions)
plot3mhistory("GB", predictions)
```

